Stochastic Variance-Reduced Policy Gradient

### 2018-02-07 svrpg.tex
changed point of view on random return: actually perfect for on-line learning

We can now state the convergence guarantees for SVRPG.
\begin{restatable}[Convergence of the {\color{red}REINFORCE-}SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
{\color{red}Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.}
Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[2]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{\zeta}{N}
		+\frac{\xi}{B},
\end{align*}
where $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
\end{restatable}
% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
Refer to App.~\ref{app:proofs} for a detailed proof involving definition of the constants and the meta-parameter constraints.
%The theoretical settings may be too conservative.
In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$ which is more suited for applications. 
By analysing the upper-bound in Thm.~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$. Advantage over batch gradient ascent can be achieved with properly selected meta-parameters. In the following we propose practical ways to do so.

>>>>

We can now state the convergence guarantees for SVRPG.
\begin{restatable}[Convergence of the {\color{red}REINFORCE-}SVRPG algorithm]{theorem}{convergence}\label{theo:convergence}
{\color{red}Consider the REINFORCE estimator in Eq.~\ref{E:policygradient.estimate}.}
Under Ass.~\ref{ass:bounded_score}, \ref{ass:REINFORCE} and \ref{ass:M2}, the parameter vector $\vtheta_A$ returned by Alg.~\ref{alg:svrpg} after $T=m\times S$ iterations has, for some positive constants $\psi,\zeta, \xi$ and for proper choice of the step size $\alpha$ and the epoch size $m$, the following property:
\begin{align*}
	&\EVV[]
	{\norm[2]{\nabla J(\vtheta_A)}^2} 
		\leq
		\frac{J(\vtheta^*)-J(\vtheta_0)}{\psi T} +
		\frac{\zeta}{N}
		+\frac{\xi}{B},
\end{align*}
where $\psi,\zeta,\xi$ depend only on $\GRADLOG,\HESSLOG,\VARRF,\VARIS,\alpha$ and $m$.
\end{restatable}
% Specific constants and meta-parameter constraints are provided in the proof of the theorem in Appendix \ref{app:proofs}.
Refer to App.~\ref{app:proofs} for a detailed proof involving definition of the constants and the meta-parameter constraints.
%The theoretical settings may be too conservative.
By analysing the upper-bound in Thm.~\ref{theo:convergence} we observe that: I) the $\mathcal{O}(\nicefrac{1}{T})$ term is coherent with results on \textcolor{red}{non-concave} SVRG~\citep[\eg][]{reddi2016stochastic}; II) the $\mathcal{O}(\nicefrac{1}{N})$ term is due to the FG approximation and is analogous to the one in~\citep{harikandeh2015stopwasting}; III) the $\mathcal{O}(\nicefrac{1}{B})$ term is due to importance weighting.
To achieve true convergence, both the batch size $N$ and the mini-batch size $B$ should increase over time. 
In practice, it is enough to choose $N$ and $B$ large enough to make the second and the third term negligible, \ie to mitigate the variance introduced by FG approximation and importance sampling, respectively.
Once the last two terms can be neglected, the number of trajectories needed to achieve $\norm[2]{\gradJ{\vtheta}}^2\leq\epsilon$ is $O(\frac{B+\nicefrac{N}{m}}{\epsilon})$. In this sense, advantage over batch gradient ascent can be achieved with properly selected meta-parameters. In Sec.~\ref{sec:stopping} we propose a joint selection of step size $\alpha$ and epoch size $m$.
Finally, note from the return statement of Algorithm \ref{alg:svrpg} that $J(\vtheta_A)$ can be seen as the average performance of all the policies tried by the algorithm. This is particularly meaningful in the context of on-line learning that we are addressing in this paper.
